---
title: "Module 4"
author: "Sankalp Biswal"
date: "2023-12-02"
output: html_document
---

## 1.Importing Libraries

```{r}
library(MASS)
library(tidyverse)
library(janitor)
library(knitr)
library(car)

```

## 2. Importing Dataset

```{r}
dataset <- cats
male <- subset(cats, subset = (cats$Sex=="M"))$Bwt
female <- subset(cats, subset = (cats$Sex=="F"))$Bwt
```

## 3. Testing

## Overview

-   There are 2 parts namely - **PART-1 and PART-2**.
-   **PART-1** deals with Two sample t-test with unequal variance for **cats** dataset.
-   **PART-2** deals with Paired t-test for the provided data for two significance levels, **alpha = 0.05** and **alpha = 0.1**.

------------------------------------------------------------------------

### **PART-1**

### Two sample t-test for cats dataset

-   We'll try to answer the question - do male and female cat samples have the same body weight ("Bwt")?
-   First, lets check if the assumptions for a Welch's t-test are satisfied or not.

#### a. Stating the hypothesis

1.  **Null Hypothesis (H0)** - The male and female cats have same body weight.
2.  **Alternative Hypothesis (H1)** - The male and female cats don't have same body weight

#### b. Assumptions of Welch's t-test.

1.  **Independent Samples**: The observations in each group are independent of each other.
2.  **Interval or Ratio Data**: The data being compared are measured on an interval.
3.  **Approximate Normality**: The sample size is large enough for the model to be robust against this assumption, but we'll check for normality below.
4.  **Unequal Variances**: The variances of the two groups being compared are not necessarily equal.

```{r}
par(mfrow = c(1, 2))

# Plot the first graph
hist(male)
# Plot the second graph
hist(female)

par(mfrow = c(1, 1))

```

-   Histogram for bodyweight of male cats displays somewhat normal distribution.
-   Histogram for bodyweight of female cats displays non-normal distribution.

To confirm the above, we'll conduct **Shapiro-Wilk normality test**

```{r}
shapiro.test(male)

shapiro.test(female)

```

-   **Male data:** Since p-value \> 0.05(significance level) we can conclude that this is normally distributed.

-   **Female data:** Since p-value \< 0.05(significance level) we can conclude that this is not normally distributed.

**But since our sample size is \> 30, our Welch t-test is robust to such violations**.

### c. Conducting the Welch Two Sample t - test

```{r}
t_test_result <- t.test(male, female, var.equal = FALSE)
options(scipen = 999)
print(t_test_result)

```

#### d. Interpretation

-   Criterias for rejecting null hypothesis using p-value

1.  If p-value \<= 0.05 (Standard significance value), Reject Null Hypothesis
2.  If p-value \> 0.05 (Standard significance value), Fail to reject Null Hypothesis

-   p-value: We got p-value = 0.000000000000008831 i.e p \< 0.05, therefore we **reject null hypothesis**

Confidence Interval: The 95% confidence interval for the difference in means (0.418 to 0.663) does not include zero, supporting our decision of **rejecting null hypothesis**

#### e. Decision

-   We therefore conclude that there is **significant difference between bodyweights** of male cats and female cats.

------------------------------------------------------------------------

### **PART-2**

-   We'll conduct two-sample paired t-test for alpha = 0.05 and alpha = 0.1 to see if there's difference in conclusion

### 1. Paired t-test for alpha = 0.05

#### a. Stating the hypotheses:

1.  **Null Hypothesis (H0)**: There is no difference in sleep quality before and after the meditation workshop.

-   **H0 : μ before − μ after = 0**

2.  **Alternative Hypothesis (H1)**: Meditation improves sleep quality.

-   **H1 : μ after−μ before \> 0 (One-tailed test as we are checking for improvement)**

```{r}
before <- c(4.6, 7.8, 9.1, 5.6, 6.9, 8.5, 5.3, 7.1, 3.2, 4.4)
after <- c(6.6, 7.7, 9.0, 6.2, 7.8, 8.3, 5.9, 6.5, 5.8, 4.9)

# Paired t-test
result <- t.test(before, after, paired = TRUE, alternative = "greater")

print(result)
```

#### b. Interpretation

-   Criterias for rejecting null hypothesis using p-value

1.  If p-value \<= 0.05 (Standard significance value), Reject Null Hypothesis
2.  If p-value \> 0.05 (Standard significance value), Fail to reject Null Hypothesis

-   p-value: We got p-value = 0.09584 i.e p \< 0.05, therefore we **do not reject null hypothesis**
-   Confidence interval: The confidence interval provides a range of values within which we can be 95% confident the true population mean lies. Our 95 percent confidence interval is **-1.203405 - Inf** which includes **zero**. This further strengthens our conclusion of **not rejecting Null hypothesis**

#### c. Decision

-   At a 5% significance level or a 95 percent confidence level, **There is sufficient evidence to suggest that meditation does not improve sleep quality**.

------------------------------------------------------------------------

### 2. Paired t-test for alpha = 0.1

#### a. Stating the hypotheses:

1.  **Null Hypothesis (H0)**: There is no difference in sleep quality before and after the meditation workshop.

-   **H0 : μ before − μ after = 0**

2.  **Alternative Hypothesis (H1)**: Meditation improves sleep quality.

-   **H1 : μ after−μ before \> 0 (One-tailed test as we are checking for improvement)**

```{r}
before <- c(4.6, 7.8, 9.1, 5.6, 6.9, 8.5, 5.3, 7.1, 3.2, 4.4)
after <- c(6.6, 7.7, 9.0, 6.2, 7.8, 8.3, 5.9, 6.5, 5.8, 4.9)

# Paired t-test
result <- t.test(before, after, paired = TRUE, alternative = "greater", alpha = 0.1)

print(result)
```

#### b. Interpretation

-   Criterias for rejecting null hypothesis using p-value

1.  If p-value \<= 0.05 (Standard significance value), Reject Null Hypothesis
2.  If p-value \> 0.05 (Standard significance value), Fail to reject Null Hypothesis

-   p-value: We got p-value = 0.9584 i.e p \< 0.05, therefore we **do not reject null hypothesis**
-   Confidence interval: The confidence interval provides a range of values within which we can be 95% confident the true population mean lies. Our 95 percent confidence interval is **-1.203405 - Inf** which includes **zero**. This further strengthens our conclusion of **not rejecting Null hypothesis**

#### c. Decision

-   At a 5% significance level or a 95 percent confidence level, **There is sufficient evidence to suggest that meditation does not improve sleep quality**.

-   Therefore, our decision doesn't change whether we take alpha = 0.05 or alpha = 0.1.

------------------------------------------------------------------------

### Justification:

1.  **Choice of Test**: We use a paired t-test because we are comparing the same group before and after an intervention. This is appropriate for assessing the impact of the meditation workshop on sleep quality for each individual.

2.  **Why Paired T-Test**: Other tests like the Two-Sample t-Test wouldn't be suitable here because we are not comparing two independent groups; instead, we are comparing measurements from the same individuals before and after treatment. The paired t-test accounts for individual variability and is more sensitive to changes within the same group.

------------------------------------------------------------------------
